Direct observation of Feshbach enhanced \swave scattering of fermions

### 2015-08-15 draft1.tex
After second round of Lauren and Hsin-I comments, sent to Ian.

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave }
\newcommand{\pwave}[0]{$\it{p}$-wave }
\newcommand{\K}{$^{40}\rm{K}$}
\newcommand{\Rb}{$^{87}\rm{Rb}$}
\newcommand{\us}{$\rm{\mu s}$} 
\newcommand{\mT}{$\rm{mT}$}
\newcommand{\ez}{$\bf{\hat{e}_z}$}
\newcommand{\ex}{$\bf{\hat{e}_x}$}
\newcommand{\um}{$\rm{\mu m}$}

@@

Feshbach resonances are widely used for tuning the interaction strength in ultracold atomic gases. They have been particularly instrumental in the study of interactions and interaction-dependent processes in cold Fermi gases. In contrast to atomic Bose-Einstein condensates (BECs), where even weak interactions play a crucial role, for example giving rise to their characteristic Thomas-Fermi density profiles \cite{KetterleBEC}, the effects of interactions in Fermi gases are harder to observe. The density of Fermi clouds is typically $\sim1000$ times less than that of BECs, making it necessary to enhance the strength of interactions in order to observe them \cite{KetterleDFG}. The tunability of interactions provided by Feshbach resonances has allowed for creation of molecular Bose-Einstein condensates from Fermi gases \cite{Greiner03,Zwierlein03, Jochim03} as well as observation of the phase transition from the BCS to BEC regime at sufficiently low temperatures \cite{Bartenstein04, Bourdel04, Zwierlein04, Regal04}. 
\par A Feshbach resonance occurs when a two-atom bound molecular state energetically approaches the two-atom continuum \cite{Chin10, Timmermans99}. In experiment, the relative energy of the free atomic states and the molecular state is defined by a bias magnetic field, so the Feshbach resonance can be approached by changing the bias field. In the simple case where there are no inelastic two-body channels, such as for the \K{} resonance discussed in this work, the effect of the resonance on the scattering length between two free atoms is \cite{Chin10}

@@

 $a_{bg}$ is the bachground scattering length, $\Delta$ is the width of the resonance, and $B_0$ is the field value at which the resonance occurs.The scattering length diverges at the resonance.

@@

\par In our dilute Fermi clouds, even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms scattered as the clouds passed through each other. This made direct detection of \swave scattering halos difficult due to detection uncertainty, that disproportionately affected regions of low atomic density in absorption images. To optimize the signal to noise for low atom numbers, we absorption imaged with fairly long, high-intensity pulses --- a non-standard regime, where the atoms acquired a velocity during imaging and the resulting Doppler-shift was non-negligible.  Simulation of the absorption imaging process was necessary for an accurate interpretation of these images. With the simulation corrected images, we extracted the fraction of atoms scattered in our collision experiment.

@@

Absorption imaginmeasures the shadow cast by an atomic ensemble in an illuminating probe laser beam on a camera. This imaging technique relies on optical transitions between the ground and certain excited atomic states. Such atomic transitions have an energy difference $\hbar\omega_0$, and a natural transition linewidth $\Gamma$. When interacting with a laser field an atom scatters photons from the laser field into the empty vacuum modes. The  rate at which this scattering occurs, in the two-level atom approximation, is \cite{LCT}

@@

\par An absorption image is obtained by shining an on- or near-resonant probe beam ($\delta\approx0$) onto the atomic cloud. Some of the light is scattered by the atoms, and the unscattered light, $I_f(x,y)$, is imaged onto a camera, as depicted in Fig. \ref{fig:absorptionIntor}a (top). The probe light is reapplied with the atoms absent to calibrate the intensity $I_0(x,y)$ of light unaffected by the atoms (bottom). 

@@

\caption{Absorption imaging. a. Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. After the atoms depart an image of just the probe light is taken. b.  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction \ez{} is lowered.  c. An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net velocity of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }  

@@

\par The intensity $I_f(x,y)$ is related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis \ez{} through a 3-D atomic density profile $\rho(x,y,z)$. For the purposes of this discussion, let us focus on a single pixel of the camera, and thus a single value of $I_0$ and $I_f$, and a single column of atoms, $\rho(z)$. We are not able to infer the entire atomic distribution, but for each pixel we may obtain the column density, $n = \int \rho\left(z\right) \,\mathrm{d}z$. As the light travels through a column of atoms, each atom scatters light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ experience a reduced optical intensity due to attenuation of the laser field by the other atoms (Fig. \ref{fig:absorptionIntor}b). The intensity change from scattering as a function of $z$ is

@@

\frac{dI(z)}{dz}=\hbar\omega_L\rho(z)\gamma_{sc}(z)=-\rho(z)\sigma_0\frac{I(z)}{1+D^2 +I/I_{\rm{sat}}},

@@

where $\sigma_0$ is the resonant scattering cross section and we define a scaled detuning $D=2\delta/\Gamma$. Integrating this equation yields  \cite{Reinaudi07}

@@

\sigma_0 n^{\rm{(1)}} = -\left(1+D^2\right)\ln\left(\frac{I_f}{I_0}\right) + \frac{I_0-I_f}{I_{\rm{sat}}}.

@@

When the probe intensity is much smaller than the saturation intensity, $I_0\ll I_{\rm{sat}}$, the right hand side of Eq. (\ref{eq2}) reduces to the optical depth, $OD=-\ln \left(I_f/I_0\right)$   \cite{Reinaudi07}. This is the simplest possible model relating the observed intensity to the column density $\sigma_0 n$. When the probe intensity is comparable to the saturation intensity, the second term in Eq. (\ref{eq2}) becomes significant, and we call this the column density corrected for high probe intensity $\sigma_0 n^{\rm{(1)}}$.
\par Equations (\ref{eq:dIdz})-(\ref{eq2}) neglect the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}. When an atom absorbs a photon from the laser light field it acquires a momentum kick $\hbar k_R$ in the direction of the light field. The associated recoil velocity is $v_r=\hbar k_R/m$, where $k_R=2\pi/\lambda$ is the recoil momentum of the light field with wavelength $\lambda$ and $m$ is the atomic mass. The re-emitted photon imparts a similar recoil momentum $\vec{p}_e$. Over many scattering events this momentum distribution averages to zero. Therefore the atom  will acquire an average per-photon velocity $v_r$ along \ez{}.  The variance of $\vec{p}_e$, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. We will ignore this correction, but the effect of this on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}c. 
\par This velocity parallel the light field Doppler shifts the atoms from resonance. After scattering $N$ photons an atom acquires an average velocity $N v_r$ and an additional detuning $\delta=k N v_r$. Therefore, even if the probe beam is initially on resonance with the atomic transition, we cannot neglect the detuning's  effect on the scattering rate. Furthermore, this detuning varies both with imaging time $t$ and with distance along the propagation direction $z$ (Fig. \ref{fig:expos}a). Thus, the intensity lost to the atoms also acquires a time dependence: 

@@

\frac{dI(t,z)}{dz}=\sigma_0 \rho(t,z) \frac{I(t,z)}{1+D(t,z)^2 +I(t,z)/I_{\rm{sat}}}, \label{eq3}

@@

As shown in  Fig. \ref{fig:expos}b, the perturbative treatment is accurate to times up to the recoil time, defined as the time it would take a single atom to be detuned by half an atomic linewidth, but diverges after that point. To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process to obtain $I_f$ as a function of imaging time, atomic density, and probe intensity. 

@@

\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms did not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_L \gamma_{sc} \rho)$. We verified this approach in known limits and then cross-checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We then developed a quasi-classical approach and allowed the atoms to move during the imaging time. While the atomic trajectories were wildly different than in the static atom approximation, the predicted $OD$s only differed on the 0.5$\%$ level.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we simulated a1-D Gaussian density distribution $\rho(z)$. We divided the cloud into spatial bins (the bin size was decreased until the result only differed at the $10^{-5}$ level from smaller bin sizes).  In this approximation, the number of atoms in each bin was time-independent.  The algorithm used is shown in Alg. [\ref{algorithm1}]. We call the optical depth simulated by this algorithm the simulated optical depth $OD^{\rm{sim1}}$

@@

\par We checked the validity of our simulation in the limits where the problem is analytically solvable. In the limit where the probe intensity is much weaker than the saturation intensity, $I_0\ll I_{\rm{sat}}$, the atoms' velocities are hardly changed, and Eq.(\ref{eq3}) reduces to 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I(z),
\end{equation}
from which we recover the analytic form 
\begin{equation}
\sigma_0 n^{\rm{(0)}} = -\ln I_0/I_f. \label{eq6}
\end{equation}
In the limit where the probe intensity is much larger than the saturation intensity, $I_0/I_{\rm{sat}}\gg D$, even far detuned atoms will scatter light at their maximum. The time dependence of the detuning can thus be neglected, and Eq. (\ref{eq3}) becomes 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I_{\rm{sat}}, 
\end{equation}
which integrates to 
\begin{equation}
\sigma_0 n = \frac{I_0 - I_f}{I_{\rm{sat}}}. \label{eq8}
\end{equation}
We recognize the right hand sides of Eq. (\ref{eq6}) and Eq. (\ref{eq8}) as the two terms in Eq. (\ref{eq2}). Thus, as shown in  Fig. \ref{fig:IsatLimits}, $OD^{\rm{sim1}}$  coincides with the optical depth as predicted by Eq. (\ref{eq2}) in both the small and large probe intensity limits. 
\begin{figure}
	\includegraphics{figure3.pdf}
\caption{Optical depth as a function of probe intensity as predicted by the simulation (blue symbols) and by Eq. (\ref{eq2}) (green curves), for three different imaging times. The predictions agree in both the high and low intensity limits, and differ for probe intensities comparable to the saturation intensity and longer imaging times. }  
\label{fig:IsatLimits}
\end{figure}
\par We used the results of this simulation to cross-check if the validity of the  stationary atom assumption, i.e. the distance traveled by the atoms (as deduced from integrating the acquired recoil velocity over the imaging time) is less than the bin size. As can be seen from Fig. \ref{fig:simTests}a, not only do the atoms travel more than the bin size, but they travel far beyond the initial extent of the cloud. Moreover, owing to the higher initial scatter rate, the back of the cloud overtakes the front for long imaging times. Thus, the atomic distribution as a function of position changes dramatically during the imaging pulse, and the stationary assumption is invalid. 
\begin{figure}
	\subfigure[]{\includegraphics{figure4.pdf}}
	\subfigure[]{\includegraphics{figure5.pdf}}
\caption{(a) Position of atoms as a function of imaging time for atoms in the first (solid green), middle (dashed red), and last (dotted blue) bins of the simulated density distribution for an initial cloud 50 \um{} in extent. The probe intensity used in this calculation was $1.2\, I_{\rm{sat}}$, and the column density was $\sigma_0 n=1.6$. (b) The velocity of a single composite atom as a function of probe intensity for various imaging times. Simulation data (dots) and numerical  solutions of Eq. (\ref{eq10}) (lines) are in agreement.}  
\label{fig:simTests}
\end{figure}

@@

To account for the changing atomic distribution during the imaging pulse, we numerically simulated the classical kinetics of atoms subject to the recoil driven optical forces. To simulate large ensembles in a reasonable time, we modeled composite `super-atoms', each describing the aggregate behavior of $N_{ca}$ atoms. The amended algorithm is shown in Alg.[\ref{algorithm2}]. 

@@

\STATE $z[n]=z_0$, $\delta[n]=0$ \COMMENT{initialize position and detuning for each composite atom, labeled by index $n$}
\STATE $O[i]=n$ \COMMENT{make a list of composite atom indexes, ordered by position}

@@

\STATE $n=O[i]$ \COMMENT{apply probe intensity to composite atoms in order of appearance}
 \STATE $A=\sigma_0 N_{sa} dz$ \COMMENT{dz is length over which atoms were grouped into single composite atom}

@@

\STATE $O[i]$=sort($n$, key =$z[n]$) \COMMENT{sort composite atom indexes by current position}

@@

\par To validate our code, we again checked the velocity predicted in this model against known limits. One such limit is that of a single composite atom. In this case, there is no attenuation, and the intensity seen by the composite atom is constant at $I_0$. Only the detuning  evolves in time, and Eqs. (\ref{eq3}) and (\ref{eq4}) give

@@

where $D = 2\delta/\Gamma$, and $\tilde{I} = I_0/I_{\rm{sat}}$. Equation (\ref{eq10}) can be solved numerically, and is in agreement with our simulation, as seen in Fig. \ref{fig:simTests}b.

@@

\par We compared the optical depths predicted by each of the two models, $OD^{\rm{sim1}}$ and $OD^{\rm{sim2}}$. As seen Fig. \ref{fig:compareModelsAndIsat}a, the predicted optical depths were hardly changed by including the full time evolution:  $\left|OD^{\rm{sim1}}-OD^{\rm{sim2}}\right|/OD^{\rm{sim1}} \le 0.005$. Thus, for the purposes of deducing the atom density from experimental optical depths, the stationary atom model is sufficient. Furthermore, we simulated a range of initial density profiles $\rho(z)$, and found their impact to be negligible \--- the only observable is the integrated atomic density $n=\int\rho(z)\mathrm{d}z$. For interpreting experimental images, we used the data generated by the travelling atom simulation.

@@

\caption{(a) Top. Optical depth as a function of probe intensity for an imaging time $t=100$ \us. The two versions of simulated optical depth, $OD^{\rm{sim1}}$ (green curve) and $OD^{\rm{sim2}}$ (green dots) are plotted. Bottom. The percent difference, $100*\left|OD^{\rm{sim1}}-OD^{\rm{sim2}}\right|/OD^{\rm{sim1}} $, from the top graph. (b) The optical depth as a function of probe intensity for three imaging times: $t=40$\us{} (blue),  $t=75$\us{} (green),  $t=100$\us{} (red). The dots represent experimental data and the lines represent the best fit of simulated data. The optimal fit parameters pictured are a $\sigma_0 n$ of 1.627(5) and saturation intensity of 29(7) counts/\us{}.  }  

@@

\subsection{Signal to noise optimization}
This simulation allowed us to interpret experimental data. For a given imaging time, we created a look-up table of predicted optical depth as a function of probe intensity and atomic column density. We then found the observed optical depth on this table, with the given probe intensity, and inferred the atomic density. The uncertainty in the measured intensities can be propagated through this procedure, and we established optimal imaging parameters to maximize the signal to noise ratio of this detection scheme. 
\par The only source of measurement uncertainty we considered was the Poisson distributed photon shot noise, with a standard deviation proportional to $\sqrt{N_p}$, where $N_p$ is the photon number. We then propagated this uncertainty through our correction scheme to obtain the uncertainty in our deduced value of $\sigma_0 n$. We define the signal to noise ratio (SNR) as $\sigma_0 n/\delta_{\sigma_0 n}$, where $ \delta_{\sigma_0 n}$ is the propagated measurement uncertainty.
\par As seen in Fig. \ref{fig:SNR}a, after about 40 \us{} extending the imaging time no longer yields appreciable improvement in SNR. Imaging for 40 \us{} as opposed to 10 \us{} where the uncorrected model is appropriate, improves the SNR by a factor of  1.5. We performed the experiments described in the second section at 40 \us{} imaging time. Figure \ref{fig:SNR}b shows that the optimal probe intensity varies with the atomic column density. For low atom numbers, $\sigma_0 n\approx0.1$, a probe intensity of $I_0\approx0.6\,I_{\rm{sat}}$ is best. However, in our experiment the probe intensity had a gaussian profile and was not uniform over the whole image.  The typical probe intensities used in our experiments varied over the $I_0=0.1\,I_{\rm{sat}}-0.7\,I_{\rm{sat}}$  range.
\begin{figure}
	\subfigure[]{\includegraphics{figure9a.pdf}}
	\subfigure[]{\includegraphics{figure9b.pdf}}
\caption{Signal to noise ratio (SNR) for three different column densities after correcting for recoil induced detuning. (a) SNR as a function of imaging time fora probe intensity of $I_0/I_{\rm{sat}}=5.0$ and (b) SNR as a function of probe intensity for an imaging time of 50 \us{}.}  
\label{fig:SNR}
\end{figure}

>>>>

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave}
\newcommand{\pwave}[0]{$\it{p}$-wave}
\newcommand{\K}{$^{40}\rm{K}$}
\newcommand{\Rb}{$^{87}\rm{Rb}$}
\newcommand{\us}{$\rm{\mu s}$} 
\newcommand{\mT}{$\rm{mT}$}
\newcommand{\ez}{$\bf{\hat{e}_z}$}
\newcommand{\ex}{$\bf{\hat{e}_x}$}
\newcommand{\um}{$\rm{\mu m}$}

@@

Feshbach resonances are widely used for tuning the interaction strength in ultracold atomic gases. They have been particularly instrumental in the study of interactions and interaction-dependent processes in cold Fermi gases. In contrast to atomic Bose-Einstein condensates (BECs), where even weak interactions play a crucial role, for example giving rise to their characteristic Thomas-Fermi density profiles \cite{KetterleBEC}, the effects of interactions in Fermi gases are harder to observe. The density of Fermi clouds is typically $\sim1000$ times less than that of BECs, making it necessary to enhance the strength of interactions in order to observe significant interaction effects\cite{KetterleDFG}. The tunability of interactions provided by Feshbach resonances has allowed for creation of molecular Bose-Einstein condensates from Fermi gases \cite{Greiner03,Zwierlein03, Jochim03} as well as observation of the phase transition from the Bardeen-Cooper-Schrieffer (BCS) superconduting regime to the BEC regime at sufficiently low temperatures \cite{Bartenstein04, Bourdel04, Zwierlein04, Regal04}. 
\par A Feshbach resonance occurs when the two-atom continuum energetically approaches a two-atom bound molecular state  \cite{Chin10, Timmermans99}. In experiment, the relative energy of the free atomic states in two hyperfine sublevels and the molecular state is defined by a bias magnetic field, so the Feshbach resonance can be approached by changing the bias field. In the simple case where there are no inelastic two-body channels, such as for the \K{} resonance discussed in this work, the effect of the resonance on the scattering length between two free atoms is \cite{Chin10}

@@

 $a_{bg}$ is the bachground scattering length, $\Delta$ is the width of the resonance, and $B_0$ is the field value at which the resonance occurs. The scattering length diverges at the resonance.

@@

\par In our dilute DFGs, even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms scattered as the clouds passed through each other. This made direct detection of \swave scattering halos difficult due to detection uncertainty, that disproportionately affected regions of low atomic density in absorption images. To optimize the signal to noise for low atom numbers, we absorption imaged with fairly long, high-intensity pulses --- a non-standard regime, where the atoms acquired a velocity during imaging and the resulting Doppler-shift was non-negligible.  Simulation of the absorption imaging process was necessary for an accurate interpretation of these images. With the simulation corrected images, we extracted the fraction of atoms scattered in our collision experiment.

@@

Absorption imaging measures the shadow cast by an atomic ensemble in an illuminating probe laser beam on a camera. This imaging technique relies on optical transitions between the ground and certain excited atomic states. Such atomic transitions have an energy difference $\hbar\omega_0$, and a natural transition linewidth $\Gamma$. When interacting with a laser field an atom scatters photons from the laser field into the vacuum modes. The  rate at which this scattering occurs, in the two-level atom approximation, is \cite{LCT}

@@

\par An absorption image is obtained by shining an on- or near-resonant probe beam ($\delta\approx0$) onto the atomic cloud. Some of the light is scattered by the atoms, and the unscattered light, $I_f(x,y)$, is imaged onto a camera, as depicted in Fig. \ref{fig:absorptionIntor}(a) (top). The probe light is reapplied with the atoms absent to calibrate the intensity $I_0(x,y)$ of light unaffected by the atoms (bottom). 

@@

\caption{Absorption imaging. (a) Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. After the atoms depart an image of just the probe light is taken. (b)  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction \ez{} is lowered.  (c) An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net velocity of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }  

@@

\par The intensity $I_f(x,y)$ is related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis \ez{} through a 3-D atomic density profile $\rho(x,y,z)$. For the purposes of this discussion, let us focus on a single pixel of the camera, and thus a single value of $I_0$ and $I_f$, and a single column of atoms, $\rho(z)$. We are not able to infer the entire atomic distribution, but for each pixel we may obtain the column density, $n = \int \rho\left(z\right) \,\mathrm{d}z$. As the light travels through a column of atoms, each atom scatters light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ experience a reduced optical intensity due to attenuation of the laser field by the other atoms (Fig. \ref{fig:absorptionIntor}(b)). The intensity change from scattering as a function of $z$ is

@@

\frac{dI(z)}{dz}=-\hbar\omega_L\rho(z)\gamma_{sc}(z)=-\rho(z)\sigma_0\frac{I(z)}{1+D^2 +I/I_{\rm{sat}}},

@@

where $\sigma_0$ is the resonant scattering cross section and we define a scaled detuning $D=2\delta/\Gamma$. Integrating this equation yields a relation between the observed optical depth $OD=-\ln \left(I_f/I_0\right)$ and the atomic column density $n$ \cite{Reinaudi07} --- we call the column density deduced in this manner $\sigma_0 n^{\rm{(1)}}$:

@@

\sigma_0 n^{\rm{(1)}} = \left(1+D^2\right)OD+ \frac{I_0-I_f}{I_{\rm{sat}}}.

@@

When the probe intensity is much smaller than the saturation intensity, $I_0\ll I_{\rm{sat}}$, and the probe light is on resonance, $D=0$, the right hand side of Eq. (\ref{eq2}) reduces to the optical depth \cite{Reinaudi07}. The column density obtained with this simplest relation is $\sigma_0 n^{\rm{(0)}}=OD$. When the probe intensity is comparable to the saturation intensity, the second term in Eq. (\ref{eq2}) becomes significant, and we call this the column density corrected for high probe intensity, $\sigma_0 n^{\rm{(1)}}$.
\par Equations (\ref{eq:dIdz})-(\ref{eq2}) neglect the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}. When an atom absorbs a photon from the laser light field it acquires a momentum kick $\hbar k_R$ in the direction of the light field. The associated recoil velocity is $v_r=\hbar k_R/m$, where $\hbar k_R= h/\lambda$ is the recoil momentum of the light field with wavelength $\lambda$ and $m$ is the atomic mass. The re-emitted photon imparts a similar recoil momentum $\vec{p}_e$, but over many scattering events this momentum distribution averages to zero. Therefore the atom  will acquire an average per-photon velocity $v_r$ along \ez{}.  The variance of $\vec{p}_e$, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. We will ignore this correction, but the effect of this on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}(c). 
\par This velocity parallel the light field Doppler shifts the atoms from resonance. After scattering $N$ photons an atom acquires an average velocity $N v_r$ and an additional detuning $\delta=k_R N v_r$. Therefore, even if the probe beam is initially on resonance with the atomic transition, we cannot neglect the detuning's  effect on the scattering rate. Furthermore, this detuning varies both with imaging time $t$ and with distance along the propagation direction $z$ (Fig. \ref{fig:expos}(a)). Thus, the intensity lost to the atoms also acquires a time dependence: 

@@

\frac{dI(t,z)}{dz}=-\sigma_0 \rho(t,z) \frac{I(t,z)}{1+D(t,z)^2 +I(t,z)/I_{\rm{sat}}}, \label{eq3}

@@

As shown in  Fig. \ref{fig:expos}(b), the perturbative treatment is accurate to times up to the recoil time, defined as the time it would take a single atom to be detuned by half an atomic linewidth, but diverges after that point. To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process to obtain $I_f$ as a function of imaging time, atomic density, and probe intensity. 

@@

\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms did not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_L \gamma_{sc} \rho)$. We verified this approach in known limits and then cross-checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We then developed a quasi-classical approach and allowed the atoms to move during the imaging time. While the atomic trajectories were wildly different than in the static atom approximation, we found that the predicted $OD$s only differed on the 0.5$\%$ level.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we simulated a 1-D Gaussian density distribution $\rho(z)$. We divided the cloud into spatial bins (the bin size was decreased until the result only differed at the $10^{-5}$ level from smaller bin sizes).  In this approximation, the number of atoms in each bin was time-independent.  The algorithm used is shown in Alg. (\ref{algorithm1}). We call the optical depth simulated by this algorithm the simulated optical depth $OD^{\rm{sim1}}$

@@

\par We checked the validity of our simulation in the limits where the problem is analytically solvable. In the limit where the probe intensity is much weaker than the saturation intensity, $I_0\ll I_{\rm{sat}}$, the atoms' velocities are hardly changed, and Eq.(\ref{eq3}) reduces to 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I(z),
\end{equation}
from which we recover the analytic form 
\begin{equation}
\sigma_0 n^{\rm{(0)}} = -\ln I_0/I_f. \label{eq6}
\end{equation}
In the limit where the probe intensity is much larger than the saturation intensity, $I_0/I_{\rm{sat}}\gg D$, even far detuned atoms will scatter light at their maximum. The time dependence of the detuning can thus be neglected, and Eq. (\ref{eq3}) becomes 
\begin{equation}
\frac{dI(z)}{dz}=-\rho\sigma_0 I_{\rm{sat}}, 
\end{equation}
which integrates to 
\begin{equation}
\sigma_0 n = \frac{I_0 - I_f}{I_{\rm{sat}}}. \label{eq8}
\end{equation}
We recognize the right hand sides of Eq. (\ref{eq6}) and Eq. (\ref{eq8}) as the two terms in Eq. (\ref{eq2}). Thus, as shown in  Fig. \ref{fig:IsatLimits}, $OD^{\rm{sim1}}$  coincides with the optical depth as predicted by Eq. (\ref{eq2}) in both the small and large probe intensity limits. 
\begin{figure}
	\includegraphics{figure3.pdf}
\caption{Optical depth as a function of probe intensity as predicted by the simulation (blue symbols) and by Eq. (\ref{eq2}) (green curves), for three different imaging times. The predictions agree in both the high and low intensity limits, and differ for probe intensities comparable to the saturation intensity and longer imaging times. }  
\label{fig:IsatLimits}
\end{figure}
\par We used the results of this simulation to cross-check if the validity of the  stationary atom assumption, i.e. the distance traveled by the atoms (as deduced from integrating the acquired recoil velocity over the imaging time) is less than the bin size. As can be seen from Fig. \ref{fig:simTests}(a), not only do the atoms travel more than the bin size, but they travel far beyond the initial extent of the cloud. Moreover, owing to the higher initial scatter rate, the back of the cloud overtakes the front for long imaging times. Thus, the atomic distribution as a function of position changes dramatically during the imaging pulse, and the stationary assumption is invalid. 
\begin{figure}
	\subfigure[]{\includegraphics{figure4.pdf}}
	\subfigure[]{\includegraphics{figure5.pdf}}
\caption{(a) Position of atoms as a function of imaging time for atoms in the first (solid green), middle (dashed red), and last (dotted blue) bins of the simulated density distribution for an initial cloud 50 \um{} in extent. The probe intensity used in this calculation was $1.2\, I_{\rm{sat}}$, and the column density was $\sigma_0 n=1.6$. (b) The velocity of a single composite atom as a function of probe intensity for various imaging times. Simulation data (dots) and numerical  solutions of Eq. (\ref{eq10}) (lines) are in agreement.}  
\label{fig:simTests}
\end{figure}

@@

To account for the changing atomic distribution during the imaging pulse, we numerically simulated the classical kinetics of atoms subject to the recoil driven optical forces. To simulate large ensembles in a reasonable time, we modeled composite `super-atoms', each describing the aggregate behavior of $N_{ca}$ atoms. The amended algorithm is shown in Alg. (\ref{algorithm2}). 

@@

\STATE $z[n]=z_0$, $\delta[n]=0$ \COMMENT{initialize position and detuning for each superatom, labeled by index $n$}
\STATE $O[i]=n$ \COMMENT{make a list of superatom indexes, ordered by position}

@@

\STATE $n=O[i]$ \COMMENT{apply probe intensity to superatoms in order of appearance}
 \STATE $A=\sigma_0 N_{sa} dz$ \COMMENT{dz is length over which atoms were grouped into single superatom}

@@

\STATE $O[i]$=sort($n$, key =$z[n]$) \COMMENT{sort superatom indexes by current position}

@@

\par To validate our code, we again checked the velocity predicted in this model against known limits. One such limit is that of a single `superatom'. In this case, there is no attenuation, and the intensity seen by the `superatom' is constant at $I_0$. Only the detuning  evolves in time, and Eqs. (\ref{eq3}) and (\ref{eq4}) give

@@

where $D = 2\delta/\Gamma$, and $\tilde{I} = I_0/I_{\rm{sat}}$. Equation (\ref{eq10}) can be solved numerically, and is in agreement with our simulation, as seen in Fig. \ref{fig:simTests}(b).

@@

\par We compared the optical depths predicted by each of the two models, $OD^{\rm{sim1}}$ and $OD^{\rm{sim2}}$. As seen Fig. \ref{fig:compareModelsAndIsat}(a), the predicted optical depths were hardly changed by including the full time evolution:  $\left|OD^{\rm{sim1}}-OD^{\rm{sim2}}\right|/OD^{\rm{sim1}} \le 0.005$. Thus, for the purposes of deducing the atom density from experimental optical depths, the stationary atom model is sufficient. Furthermore, we simulated a range of initial density profiles $\rho(z)$, and found their impact to be negligible \--- the only observable is the integrated atomic density $n=\int\rho(z)\mathrm{d}z$. For interpreting experimental images, we used the data generated by the travelling atom simulation.

@@

\caption{(a) Top. Optical depth as a function of probe intensity for an imaging time $t=100$ \us. $OD^{\rm{(1)}}$ and $OD^{\rm{(2)}}$ are optical depths predicted from a given column density by Eq. (\ref{eq2}) and (\ref{eq:OD2}) respectively.  The two versions of simulated optical depth, $OD^{\rm{sim1}}$ (green curve) and $OD^{\rm{sim2}}$ (green dots) are plotted. Bottom. The percent difference, $100*\left|OD^{\rm{sim1}}-OD^{\rm{sim2}}\right|/OD^{\rm{sim1}} $, from the top graph. (b) The optical depth as a function of probe intensity for three imaging times: $t=40$\us{} (blue),  $t=75$\us{} (green),  $t=100$\us{} (red). The dots represent experimental data and the lines represent the best fit of simulated data. The optimal fit parameters pictured are a $\sigma_0 n$ of 1.627(5) and saturation intensity of 29(7) counts/\us{}.  }  

@@

\subsection{Signal to noise optimization}
This simulation allowed us to interpret experimental data. For a given imaging time, we created a look-up table of predicted optical depth as a function of probe intensity and atomic column density. We then found the observed optical depth on this table, with the given probe intensity, and inferred the atomic density. The uncertainty in the measured intensities can be propagated through this procedure, and we established optimal imaging parameters to maximize the signal to noise ratio of this detection scheme. 
\par The only source of measurement uncertainty we considered was the Poisson distributed photon shot noise, with a standard deviation proportional to $\sqrt{N_p}$, where $N_p$ is the photon number. We then propagated this uncertainty through our correction scheme to obtain the uncertainty in our deduced value of $\sigma_0 n$. We define the signal to noise ratio (SNR) as $\sigma_0 n/\delta_{\sigma_0 n}$, where $ \delta_{\sigma_0 n}$ is the propagated measurement uncertainty.
\par As seen in Fig. \ref{fig:SNR}(a), after about 40 \us{} extending the imaging time no longer yields appreciable improvement in SNR. Imaging for 40 \us{} as opposed to 10 \us{} where the uncorrected model is appropriate, improves the SNR by a factor of  1.5. We performed the experiments described in the second section at 40 \us{} imaging time. Figure \ref{fig:SNR}b shows that the optimal probe intensity varies with the atomic column density. For low atom numbers, $\sigma_0 n\approx0.1$, a probe intensity of $I_0\approx0.6\,I_{\rm{sat}}$ is best. However, in our experiment the probe intensity had a gaussian profile and was not uniform over the whole image.  The typical probe intensities used in our experiments varied over the $I_0=0.1\,I_{\rm{sat}}-0.7\,I_{\rm{sat}}$  range.
\begin{figure}
	\subfigure[]{\includegraphics{figure9a.pdf}}
	\subfigure[]{\includegraphics{figure9b.pdf}}
\caption{Signal to noise ratio (SNR) for three different column densities after correcting for recoil induced detuning. (a) SNR as a function of imaging time fora probe intensity of $I_0/I_{\rm{sat}}=5.0$ and (b) SNR as a function of probe intensity for an imaging time of 50 \us{}.}  
\label{fig:SNR}
\end{figure}
